Are we learning to build better models at the cost of better decisions?

### 2021-03-05 manuscript/manuscript.Rmd
wip

Global change issues are complex and outcomes are difficult to predict, making approaches to uncertainty a central part
of effective decision-making [@ @]. While some uncertainty is intrinsic to the underlying processes (stochasticity) or by
limits to what variables we can observe and how accurately we measure them (measurement uncertainty), much uncertainty 
is the result of our imperfect knowledge of the processes involved, as expressed by the structure and
parameters of mathematical models used to approximate those processes (model uncertainty). Unlike the other forms of
uncertainty, model uncertainty can be 
reduced by gathering additional data, which may rule out certain models or parameter values as implausible[^1]. 
When model parameters are estimated directly from available data, there is a risk that the best-fitting models may 
_overfit_ to patterns arising from by chance from the stochasticity or measurement error which do not reflect the
underlying process. Researchers have often favored simpler models which are less prone to over-fitting, sometimes
through explicit penalties such as information criteria [@], though these can be misleading [@]. With the rapid 
expansion of available ecological and environmental data [@], it is increasingly possible to rely instead on comparisons
between model predictions and data purposely excluded from model estimation (cross-validation), or better, 
comparisons between model predictions and data collected in the future (forecasting, @Clark2001; @Dietze2018]). 
Over sufficiently long timescales, comparing model forecasts to future observations will select the models with
greatest predictive accuracy.  Because model-based forecasts frequently play an important role in decision-making,
it is commonly assumed [e.g. @Walters1981; @Clark2001; @Dietze2018] that this approach for addressing model uncertainty
will also improve decision-making outcomes. Here, I illustrate that this need not be the case: it is quiet possible
that any approach with improves the forecast accuracy of the model(s) over time can simultaneously lead to worse 
decision outcomes. This example illustrates that knowing when increased forecast accuracy will or will not also improve 
decision outcomes in general remains an open problem, as is the challenge of what to do about it.

@@

In passive adaptive management, the manager chooses the action which maximizes that expected future utility. After observing the consequences of the action, the manager updates the probabilities each model is correct, typically by Bayes rule [e.g. @Walters1981; @Smith1981; @Ludwig1982]. In active adaptive management, the key difference is that the manager may chose an action which does not maximize the utility _given the current model uncertainty_, but instead, takes an action whose consequences will more quickly discriminate between the models and thus more rapidly reduce the uncertainty. After observing the outcome of this exploratory step, model probabilities are updated as before.  Under some scenarios the cost of this 'exploration' step can be made up later by actions based on the more precise knowledge gained as a result ('exploitation'). Alternative approaches such as scenario analysis can be considered approximations of this approach.  A scenario analysis does not rely on quantitative estimates of model probabilities or expected utility, but still consider possible outcomes over a suite of models (scenarios), select actions to minimize the risk or or other suitable objective, and update or eliminate scenarios deemed implausible over time [@Polasky2011].

>>>>

Global change issues are complex and outcomes are difficult to predict, making approaches to uncertainty a central part
of effective decision-making [@ @]. Reducing that uncertainty by improving models in response to more data
becomes available

While some uncertainty is intrinsic to the underlying processes (stochasticity) or by
limits to what variables we can observe and how accurately we measure them (measurement uncertainty), much uncertainty 
is the result of our imperfect knowledge of the processes involved, as expressed by the structure and
parameters of mathematical models used to approximate those processes (model uncertainty). Unlike the other forms of
uncertainty, model uncertainty can be 
reduced by gathering additional data, which may rule out certain models or parameter values as implausible[^1]. 
When model parameters are estimated directly from available data, there is a risk that the best-fitting models may 
_overfit_ to patterns arising from by chance from the stochasticity or measurement error which do not reflect the
underlying process [@Ginzburg2004]. Researchers have often favored simpler models which are less prone to over-fitting, sometimes
through explicit penalties such as information criteria [@], though these can be misleading [@]. With the rapid 
expansion of available ecological and environmental data [@], it is increasingly possible to rely instead on comparisons
between model predictions and data purposely excluded from model estimation (cross-validation), or better, 
comparisons between model predictions and data collected in the future (forecasting, @Clark2001; @Dietze2018]). 
Over sufficiently long timescales, comparing model forecasts to future observations will select the models with
greatest predictive accuracy.  Because model-based forecasts frequently play an important role in decision-making,
it is commonly assumed [e.g. @Walters1981; @Clark2001; @Dietze2018] that this approach for addressing model uncertainty
will also improve decision-making outcomes. Here, I illustrate that this need not be the case: it is quiet possible
that any approach with improves the forecast accuracy of the model(s) over time can simultaneously lead to worse 
decision outcomes. This example illustrates that knowing when increased forecast accuracy will or will not also improve 
decision outcomes in general remains an open problem, as is the challenge of what to do about it.

@@

In passive adaptive management, the manager chooses the action which maximizes that expected future utility. After observing the consequences of the action, the manager updates the posterior probabilities that each model is correct, typically by Bayes rule [e.g. @Walters1981; @Smith1981; @Ludwig1982; @Punt2016]. The decision strategy is then re-evaluated in light of this revised uncertainty.  In many real-world applications, it may be impossible or computationally prohibitive to solve the optimal control policy given all sources of uncertainty, as is typical in fisheries management practice [@Punt2016]. Management Strategy Evaluation (MSE) evaluates

In active adaptive management, the key difference is that the manager may chose an action which does not maximize the utility _given the current model uncertainty_, but instead, takes an action whose consequences will more quickly discriminate between the models and thus more rapidly reduce the uncertainty. After observing the outcome of this exploratory step, model probabilities are updated as before.  Under some scenarios the cost of this 'exploration' step can be made up later by actions based on the more precise knowledge gained as a result ('exploitation'). In practice, passive adaptive management approaches are more common, such as Management Strategy Evaluation (MSE @Punt2016) or Scenario Analysis [@Polasky2011]. These approaches all share the same adaptive aspect,

Owing to the computational complexity involved, these approaches typically evaluate a suite of proposed policies, rather than seeking to derive the optimal policy. A scenario analysis does not rely on quantitative estimates of model probabilities or expected utility, but still consider possible outcomes over a suite of models (scenarios), select actions to minimize the risk or or other suitable objective, and update or eliminate scenarios deemed implausible over time [@Polasky2011].
